Quantum Mechanics Assignment Eight

### 2016-03-02 quantum/latex/a8.tex
finish hw?

\section{Problem 2: Neutron Star}
Neutron stars are celestial objects made up almost exclusively of neutrons, or spin-1/2 fermions.\\

@@

\section{Problem 4: Three-particle system}
In this problem we practice writing the wave functions for three-particle systems of particles
with different symmetry rules:\\

@@

This is a huge magnetic field!\\

>>>>

\section{Problem 2: Neutron Star}
Neutron stars are celestial objects made up almost exclusively of neutrons, or spin-1/2 fermions.
They are incredibly massive, and without the Pauli Exclusion Principle would overcome the
inter-particle repulsions and collapse into a black hole. However, because of the PEP, no two
particles in a system can take on the same quantum state. I don't know the quantum numbers of
a neutron in a system like this, but I do know that no two neutrons in the system are allowed to
have the same spacial wave function, because this would violate the PEP. This requirement for each
particle to have a different spacial wave function translates to the particles needing to be in
different spaces, and hence not collapsing into the center and forming a black hole. Thus there
is a ``quantum force'' which opposes gravity and keeps the particles stable.\\

This is a really weird thing to think about. It is odd that the entire star is treated as a single
system, meaning each neutron must have different quantum numbers, since this translates to an
immense number of numbers. It is odd that the PEP manifests itself as a force which opposes gravity.
\\

Another rival explanation is that, due to the Heisenberg Uncertainty Principle, the position of
each particle is well known due to tight packing, so the momentum uncertainty must be very large.
This translates to the particles having a large range of momenta. This momenta could possibly be
the means by which the particles oppose the gravitational force, since the ``quantum force''
explanation seems weird to me.\\

@@

\section{Problem 4: Three-particle system}
In this problem we practice writing the wave functions for three-particle systems of particles
with different symmetry rules. Here A, B and C represent spacial states and $S_A, S_B$ and $S_C$ are
spin states. Particles 1, 2 and 3 are the three particles in the system.\\

\textbf{Distinguishable Particles}

The distinguishable particle case is simple. Because we are allowed to label particles, we say that
each particle has its own spacial and spin states A, B and C. There is no need for us to make our
wave function have equal probabilities of measuring 1, 2 and 3 in any state, we are allowed to say
particle 1 is in state A:\\

\begin{align*}
  \Psi_D &= \left(\Psi_1^A\Psi_2^B\Psi_3^C\right)\ket{S_A}_1\ket{S_B}_2\ket{S_C}_3\\
\end{align*}

\textbf{Bosons and Fermions}

The calculation for the boson and fermion states is more difficult. Our wave functions must reflect
that measuring a particle in state A cannot tell us whether this particle is particle 1, 2 or 3. To
achieve this we sum all the possible combinations of states to create a completely ambiguous
superposition state. This is the symmetric state.\\

\begin{align*}
  \Psi^{S} &=
  \Psi_1^A\Psi_2^B\Psi_3^C + \Psi_1^A\Psi_2^C\Psi_3^B\\
  &+ \Psi_1^B\Psi_2^A\Psi_3^C + \Psi_1^B\Psi_2^C\Psi_3^A\\
  &+ \Psi_1^C\Psi_2^A\Psi_3^B + \Psi_1^C\Psi_2^B\Psi_3^A\\
  \ket{S}^{S} &=
  \ket{S_A}_1\ket{S_B}_2\ket{S_C}_3 + \ket{S_A}_1\ket{S_C}_2\ket{S_B}_3\\
  &+ \ket{S_B}_1\ket{S_A}_2\ket{S_C}_3 + \ket{S_B}_1\ket{S_C}_2\ket{S_A}_3\\
  &+ \ket{S_C}_1\ket{S_A}_2\ket{S_B}_3 + \ket{S_C}_1\ket{S_B}_2\ket{S_A}_3\\
\end{align*}

We then use a Slater determinant to construct the antisymmetric spacial and spin states:\\

\begin{align*}
  \Psi^{A} &= \det
  \begin{vmatrix}
    \Psi_1^A & \Psi_2^A & \Psi_3^A\\
    \Psi_1^B & \Psi_2^B & \Psi_3^B\\
    \Psi_1^C & \Psi_2^C & \Psi_3^C\\
  \end{vmatrix}\\
  &= \Psi_1^A\Psi_2^B\Psi_3^C - \Psi_1^A\Psi_2^C\Psi_3^B\\
  &- \Psi_1^B\Psi_2^A\Psi_3^C + \Psi_1^C\Psi_2^A\Psi_3^B\\
  &+ \Psi_1^B\Psi_2^C\Psi_3^A - \Psi_1^C\Psi_2^B\Psi_3^A\\
  \ket{S}^A &= \det
  \begin{vmatrix}
    \ket{S_A}_1^A & \ket{S_A}_2 & \ket{S_A}_3\\
    \ket{S_B}_1^B & \ket{S_B}_2 & \ket{S_B}_3\\
    \ket{S_C}_1^C & \ket{S_C}_2 & \ket{S_C}_3\\
  \end{vmatrix}\\
  &= \ket{S_A}_1\ket{S_B}_2\ket{S_C}_3 - \ket{S_A}_1\ket{S_C}_2\ket{S_B}_3\\
  &- \ket{S_B}_1\ket{S_A}_2\ket{S_C}_3 + \ket{S_C}_1\ket{S_A}_2\ket{S_B}_3\\
  &+ \ket{S_B}_1\ket{S_C}_2\ket{S_A}_3 - \ket{S_C}_1\ket{S_B}_2\ket{S_A}_3\\
\end{align*}

To create our boson wave functions, we take any combination of individual wave functions that result
in an overall symmetric wave function:\\

\begin{align*}
  \Psi_B^{SS} &= \Psi_B^S\ket{S}^S\\
  &= \left(\Psi_1^A\Psi_2^B\Psi_3^C + \Psi_1^A\Psi_2^C\Psi_3^B
  + \Psi_1^B\Psi_2^A\Psi_3^C + \Psi_1^B\Psi_2^C\Psi_3^A
  + \Psi_1^C\Psi_2^A\Psi_3^B + \Psi_1^C\Psi_2^B\Psi_3^A\right)\\
  &\left(\ket{S_A}_1\ket{S_B}_2\ket{S_C}_3 + \ket{S_A}_1\ket{S_C}_2\ket{S_B}_3
  + \ket{S_B}_1\ket{S_A}_2\ket{S_C}_3 + \ket{S_B}_1\ket{S_C}_2\ket{S_A}_3
  + \ket{S_C}_1\ket{S_A}_2\ket{S_B}_3 + \ket{S_C}_1\ket{S_B}_2\ket{S_A}_3\right)
\end{align*}

\begin{align*}
  \Psi_B^{AA} &= \Psi_B^A\ket{S}^A\\
  &= \left(\Psi_1^A\Psi_2^B\Psi_3^C - \Psi_1^A\Psi_2^C\Psi_3^B
  - \Psi_1^B\Psi_2^A\Psi_3^C + \Psi_1^C\Psi_2^A\Psi_3^B
  + \Psi_1^B\Psi_2^C\Psi_3^A - \Psi_1^C\Psi_2^B\Psi_3^A\right)\\
  &\left(\ket{S_A}_1\ket{S_B}_2\ket{S_C}_3 - \ket{S_A}_1\ket{S_C}_2\ket{S_B}_3
  - \ket{S_B}_1\ket{S_A}_2\ket{S_C}_3 + \ket{S_C}_1\ket{S_A}_2\ket{S_B}_3
  + \ket{S_B}_1\ket{S_C}_2\ket{S_A}_3 - \ket{S_C}_1\ket{S_B}_2\ket{S_A}_3\right)\\
\end{align*}

To make our fermion states, we take all the possible antisymmetric overall wavefunctions:\\

\begin{align*}
  \Psi_F^{AS} &= \Psi^A\ket{S}^S\\
  &= \left(\Psi_1^A\Psi_2^B\Psi_3^C - \Psi_1^A\Psi_2^C\Psi_3^B
  - \Psi_1^B\Psi_2^A\Psi_3^C + \Psi_1^C\Psi_2^A\Psi_3^B
  + \Psi_1^B\Psi_2^C\Psi_3^A - \Psi_1^C\Psi_2^B\Psi_3^A\right)\\
  &\left(\ket{S_A}_1\ket{S_B}_2\ket{S_C}_3 + \ket{S_A}_1\ket{S_C}_2\ket{S_B}_3
  + \ket{S_B}_1\ket{S_A}_2\ket{S_C}_3 + \ket{S_B}_1\ket{S_C}_2\ket{S_A}_3
  + \ket{S_C}_1\ket{S_A}_2\ket{S_B}_3 + \ket{S_C}_1\ket{S_B}_2\ket{S_A}_3\right)\\
\end{align*}

\begin{align*}
  \Psi_F^{SA} &= \Psi^S\ket{S}^A\\
  &= \left(\Psi_1^A\Psi_2^B\Psi_3^C + \Psi_1^A\Psi_2^C\Psi_3^B
  + \Psi_1^B\Psi_2^A\Psi_3^C + \Psi_1^B\Psi_2^C\Psi_3^A
  + \Psi_1^C\Psi_2^A\Psi_3^B + \Psi_1^C\Psi_2^B\Psi_3^A\right)\\
  &\left(\ket{S_A}_1\ket{S_B}_2\ket{S_C}_3 - \ket{S_A}_1\ket{S_C}_2\ket{S_B}_3
  - \ket{S_B}_1\ket{S_A}_2\ket{S_C}_3 + \ket{S_C}_1\ket{S_A}_2\ket{S_B}_3
  + \ket{S_B}_1\ket{S_C}_2\ket{S_A}_3 - \ket{S_C}_1\ket{S_B}_2\ket{S_A}_3\right)\\
\end{align*}

\textbf{Many-electron system}
To describe a many-electron system, we use the following sets:\\

\begin{align*}
  A_{spacial} &= [\Psi^A, \Psi^B, ...]\\
  A_{spin} &= [\ket{S_A}, \ket{S_B}, ...]\\
\end{align*}

We also use function P(q), which returns all possible combinations of the elements of set q. Our
many-electron system is then represented by:\\

\begin{align*}
  \Psi_F^{AS} &= \Psi^A\ket{S}^S\\
  &= \left(\sum_p^{P(A_{spacial})}\prod_{i=0}^N p(i)_i\right)\det
  \begin{vmatrix}
    \ket{S_A}_1 & \ket{S_A}_2 & ... & \ket{S_A}_N\\
    \ket{S_B}_1 & \ket{S_B}_2 & ... & \ket{S_B}_N\\
    \vdots & \vdots & \vdots & \vdots\\
    \ket{S_N}_1 & \ket{S_N}_2 & ... & \ket{S_N}_N\\
  \end{vmatrix}
\end{align*}

\begin{align*}
  \Psi_F^{SA} &= \Psi^S\ket{S}^A\\
  &= \det
  \begin{vmatrix}
    \Psi^A_1 & \Psi^A_2 & ... & \Psi^A_N\\
    \Psi^B_1 & \Psi^B_2 & ... & \Psi^B_N\\
    \vdots & \vdots & \vdots & \vdots\\
    \Psi^N_1 & \Psi^N_2 & ... & \Psi^N_N\\
  \end{vmatrix}
  \left(\sum_p^{P(A_{spin})}\prod_{i=0}^N p(i)_i\right)
\end{align*}

This is a cool problem which made me think about how to generalize wave functions given the
Symmetrization Principle. Using Slater determinants was really helpful here, I'm not sure I could
have done the problem without them. My solution for representing many-electron systems is a bit
impractical, I'm interested in if and how such systems actually are represented.\\

@@

This is a huge magnetic field! I'm not entirely sure of it, but perhaps a very strong field is
necessary to split the orbitals to the level that they agree. It's important to understand when
the Zeeman effect can be treated as a perturbation and when it has to be actually included in
the original Hamiltonian. I said that whenever the Zeeman term is of a similar magnitude to the
fine structure correction, it is too big to include as a perturbation.\\
